---
title: "Chapter 1 - R Exercise"
output: html_notebook
---

# Vectors

```{r}
# Creating a vector
v <- c(3,1,4,1,5,9)
v

# Vector operations
sum(v)

# Vector indexing: Select items excluding 2 to 4
v[-(2:4)]

```

# Factorials and binomial coefficients

```{r}
n <- 4; k <- 2

# Calculate the factorial of n
factorial(n)

# Combinations without repetition 
choose(n,k)
```

# Sampling and simulation

```{r}
# Generate a random sample of 5 numbers from 1 to 10
n <- 10; k <- 5
sample(n,k)

# Generate a random sample with replacement
sample(n,k,replace=T)

# Generating a random permutation, this can be abreviated to sample(n)
sample(n,n)

# Generating sample of non-numeric vectors. 'letters' is an in-built procedure
sample(letters,4)

```

## Matching problem simulation
Let's show by simulation that the probability of a matching card in Example 1.6.4 is approximately 1 − 1/e when the deck is sufficiently large. Using R, we can perform the experiment a bunch of times and see how many times we encounter at least one matching card:
```{r}
n <- 100
r <- replicate(10^4,sum(sample(n)==(1:n)))
sum(r>=1)/10^4
```
## Birthday problem calculation and simulation
The following code uses prod (which gives the product of a vector) to calculate the probability of at least one birthday match in a group of 23 people:
```{r}
k <-23
1-prod(((365-k+1):365)/365)
```

